---
title: "data_analysis"
author: "Yong Lee"
date: "5/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())

# load necessary packages

library(tidyverse)
library(stargazer)
library(lmtest)
```

Data was collected and values were assigned in Excel using logical if statements. For attendance data where 1 indicates presence and 0 indicates absence, a string of 11 values were concatonated and examined for a succession of 3 or more 0's following after a 1. This was done to distinguish individuals who started attending online lectures later on in the semester and those who dropped out. Censorship was assigned on the basis of surviving students whose data for select periods of time were unavailable. The event variable is reserved for data wrangling in R.

```{r load}
# load entire dataset
# create event variable
# create outcome variable
## variables are same but needed separately later on
# convert censor variable to factor type

data_all <- read_csv("raw-data/Attendance.csv", col_types = cols()) %>%
  mutate(
    Event = 1 - Censor,
    Outcome = factor(1 - Censor,
      levels = c(0, 1),
      labels = c("Attended Lectures", "Dropped Out")
    ),
    Censor = factor(Censor,
      levels = c(0, 1),
      labels = c("Uncensored", "Censored")
    )
  )

# save dataset in raw-data subdirectory

saveRDS(data_all, "shiny_app/raw-data/data_all.rds")
```

```{r students}
# filter out members of the teaching staff
# select only variables of interest

data_students <- data_all %>%
  filter(Role == "Student") %>%
  select(Role, ID, Course, Gender, Lectures, Censor, Event, Outcome)

# save dataset in raw-data subdirectory

saveRDS(data_students, "shiny_app/raw-data/data_students.rds")
```

```{r attended}
# filter out students who never attendned lecture synchronously
# select only variables of interest

data <- data_all %>%
  filter(Role == "Student", Start != 0) %>%
  select(Role, ID, Course, Gender, Lectures, Censor, Event, Outcome)

# save dataset in raw data subdirectory

saveRDS(data, "shiny_app/raw-data/data.rds")
```

```{r pp}
# convert person-level to person-period data
# uncount duplicates observations by lecture value
# create period variable that uniquely identifies each lecture observation
# create exit variable that identifies period in which student drops out

data_pp <- data %>%
  uncount(Lectures, .remove = FALSE) %>%
  mutate(
    Period = as.factor(unlist(c(sapply(data$Lectures, function(x) seq(1, x))))),
    Exit = ifelse(Lectures == Period, Event, 0)
  )

# save dataset in raw data subdirectory

saveRDS(data_pp, "shiny_app/raw-data/data_pp.rds")
```

```{r models}
# generate logistic regression models
    
# model 1: unconditional model 
    
mod1 <- glm(Exit ~ 1, data = data_pp, family = "binomial")
data_pp <- data_pp %>% 
  mutate(pred_m1 = predict(mod1, type = "response"))

# model 2: exit on period
    
mod2 <- glm(Exit ~ Period, data = data_pp, family = "binomial")
data_pp <- data_pp %>%
  mutate(pred_m2 = predict(mod2, type = "response"))

# model 3: exit on period and course
    
mod3 <- glm(Exit ~ Period + Course, data = data_pp, family = "binomial")
data_pp <- data_pp %>%
  mutate(pred_m3 = predict(mod3, type = "response"))

# model 4: exit on period and gender
    
mod4 <- glm(Exit ~ Period + Gender, data = data_pp, family = "binomial")
data_pp <- data_pp %>%
  mutate(pred_m4 = predict(mod4, type = "response"))

# model 5: exit on period, course, and gender
    
mod5 <- glm(Exit ~ Period + Course + Gender, data = data_pp, family = "binomial")
data_pp <- data_pp %>%
  mutate(pred_m5 = predict(mod5, type = "response"))

# model 6: exit on period, course, gender, and interaction between course and gender
    
mod6 <- glm(Exit ~ Period + Course * Gender, data = data_pp, family = "binomial")
data_pp <- data_pp %>%
  mutate(pred_m6 = predict(mod6, type = "response"))

# generate taxonomy table of fitted models
    
stargazer(mod1, mod2, mod3, mod4, mod5, mod6,
          type = "text",
          dep.var.labels = "Log-Odds of Dropping Synchronous Lecture Attendance",
          covariate.labels = c("Lecture 2", "Lecture 3", "Lecture 4", "Lecture 5", "Lecture 6", "Lecture 7", "Lecture 8", "Lecture 9", "Lecture 10", "Lecture 11", "Course: S052", "Gender: Male", "Interaction: CoursexGender"))
```

```{r likelihood_test}
lrtest (mod2, mod3)
```